---
title: "p8105_hw6_waa2119"
author: "William Anderson"
date: "2022-11-28"
output: github_document
---


```{r, message = FALSE, warning = FALSE}

library(tidyverse)
library(lubridate)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 2

First we will read in the homicide data from the Washington Post

```{r, warning = FALSE}

homicide_data = 
  
  read_csv("hw6_data/homicide-data.csv") %>%
  
  janitor::clean_names() %>%
  
  mutate(
    reported_date = ymd(reported_date)) %>%
  
  mutate(victim_age = as.numeric(victim_age)) %>%
  
  mutate(city_state = str_c(city, ", ", state)) %>%
  
  mutate(unsolved = ifelse(disposition == "Closed without arrest" | disposition == "Open/No arrest", 1, 0)) %>%
  
  mutate(resolved = as.numeric(disposition == "Closed by arrest")) %>%
 
    filter(!city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO")) %>% 
  
  filter(victim_race %in% c("White", "Black")) %>%
  
  filter(victim_sex %in% c("Male", "Female")) %>%
  
  mutate(victim_race = fct_relevel(victim_race, "White"))


head(homicide_data, 20)

```

For the city of Baltimore, MD we will use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}

fit_logistic = 
  
homicide_data %>%
  
  filter(city_state %in% "Baltimore, MD") %>%
  
  select(resolved, victim_age, victim_race, victim_sex) %>%

  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) %>%
  
  broom::tidy(conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE) %>%
  
  mutate(OR = estimate) %>%
  
  mutate(log_OR = log(estimate)) %>%
  
  select(term, log_OR, OR, conf.low, conf.high, p.value)


head(fit_logistic, 20)
  
```


Now we will run the glm function for each of the cities in the dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.


```{r}
fit_all = 
  
  homicide_data %>%
  
  nest(df = -city_state) %>%
  
  mutate(
    glm_fits = map(.x = df, ~glm(resolved ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    
    results = map(glm_fits, ~broom::tidy(x = .x, conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE))) %>%
  
  unnest(results) %>% 
  
  mutate(OR = estimate) %>%
  
  mutate(log_OR = log(estimate)) %>%
  
  filter(term %in% "victim_sexMale") %>%
  
  select(city_state, term, log_OR, OR, conf.low, conf.high, p.value)

head(fit_all, 20)
```


Now we will create a plot that shows the estimated ORs and CIs for each city and organize cities according to estimated OR.

```{r}

 ggplot(fit_all, aes(x = OR, y = reorder(city_state, OR))) + 
    
  geom_point() + 
    
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +   
  
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5), legend.position = "none") + 
  
  labs(
    title = "Odds ratio of solved homicides of male victims in US cities", 
    
    y = "City, State", 
    
    x = "Odds Ratio"
  )

```


The odds of being male in a solved homicide in Albuquerque is 1.76 times more likely than being female and is the highest in the USA. The odds of being male in a solved homicide in New York City is 0.26 times more likely than being female and is the lowest in the USA. We see that in the majority of cities it is less likely to be a male victim in a solved homicide than being female as the majority of cities have an odds ratio that is less than 1 for being a male victim. Only 6 cities out of the 47 in this data set had an odds ratio greater than 1 for being a male victim in a solved homicide. However there is significant dispersion in the data and the confidence intervals are fairly large so the odds ratio data must be interpreted with caution. 


## Problem 3

First we will load the birthweight data

```{r}

birth_weight = read_csv("hw6_data/birthweight.csv", na = "") %>%
  
  mutate(babysex = as.factor(babysex),
         
         frace = as.factor(frace),
         
         malform = as.factor(malform),
         
         mrace = as.factor(mrace))
  
  anyNA(birth_weight)
  

```


I am proposing a model for birthweight with the predictors being number of cigarettes smoked by the mother during pregnancy, the presence of malformations, the babys head circumference, the baby's length, and the gestational age in weeks. I believe these factors play a significant role in determining the baby's weight at birth because the mother smoking cigarettes affects fetus development and its birth weight, malformations can affect weight causing the baby to weigh less at birth, head circumference is directly linked to weight as the larger circumference leads to higher weight, and the longer a baby is the more they will weigh at birth. Additionally, premature gestational age results in a lower weight for the child at birth. All of these

```{r}
birth_model = 
  
  lm(bwt ~ smoken + malform + bhead + blength + gaweeks, data = birth_weight)


  birth_model %>%
  
  broom::tidy() %>%
  
  select(term, estimate, p.value)
  

  birth_weight %>%
  
  modelr::add_residuals(birth_model) %>%
  
  modelr::add_predictions(birth_model) %>% 
    
  ggplot(aes(x = pred, y = resid)) +
  
  geom_point(alpha = 0.3) + 
  
  geom_smooth(se = FALSE, method = "lm") + 
  
  labs(title = "Residuals vs Fitted values for my model", 
       x = "Fitted Values", 
       
       y = "Residuals")
  
  
```

Looking at the residuals vs fitted values for the model I developed we see that the residual values bounce around the zero line which suggests that a linear relationships between the terms is appropriate to model. 

Now we will compare this model to two others:

* One using length at birth and gestational age as predictors (main effects only)

* One using head circumference, length, sex, and all interactions (including the three-way interaction). 

```{r}
compare1 = lm(bwt ~ blength + gaweeks, data = birth_weight)

compare2 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + babysex*blength + bhead*blength*babysex, data = birth_weight)


cv_df = crossv_mc(birth_weight, 100)

cv_df = 
  
  cv_df %>%
  
  mutate(
    train = map(train, as_tibble),
    
    test = map(test, as_tibble))

cv_compare = 
  cv_df %>%
  
  mutate(
    my_model = map(train, ~lm(bwt ~ smoken + malform + bhead + blength + gaweeks, data = .x)), 
    
    length_gaweeks = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    
    head_length_sex = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + babysex*blength + bhead*blength*babysex, data = .x))) %>%
  
  mutate(
    rmse_my_model = map2_dbl(my_model, test, ~rmse(model = .x, data = .y)),
    
    rmse_length_gaweeks = map2_dbl(length_gaweeks, test, ~rmse(model = .x, data = .y)),
    
    rmse_head_length_sex = map2_dbl(head_length_sex, test, ~rmse(model = .x, data = .y)))

cv_compare %>%
  
  select(starts_with("rmse")) %>%
  
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  
  mutate(model = fct_inorder(model)) %>%
  
  ggplot(aes(x = model, y = rmse)) + 
  
  geom_violin()
```

We see there is higher error in the model that utilizes the baby's length and gestational age as predictors for birth weight. The model I developed which has predictors of predictors being number of cigarettes smoked by the mother during pregnancy, the presence of malformations, the babys head circumference, the baby's length, and the gestational age, and the model containing the baby's head circumference, length, and sex have comparable RMSE values but there is more dispersion in the model I developed. 
